## Question - 1

```{r}
library(tidyverse)
library(caret)
Model_Building_Data <- read.csv("Model_Building_Data.csv")
str(Model_Building_Data)
summary(Model_Building_Data)

```
## Observations : 
- There are 6488 observations and 17 variables.

- Most variables are categorical.

- TotalCharges has some NA values.

```{r}
# Checking for missing values
colSums(is.na(Model_Building_Data))

# Removing rows with missing TotalCharges
Model_Building_Data <- Model_Building_Data[!is.na(Model_Building_Data$TotalCharges), ]

```


```{r}
# Visualize tenure
ggplot(Model_Building_Data, aes(x = Tenure, fill = Status)) +
  geom_histogram(binwidth = 5, position = "dodge") +
  labs(title = "Tenure Distribution by Customer Status")

# Visualize Monthly Charges
ggplot(Model_Building_Data, aes(x = MonthlyCharges, fill = Status)) +
  geom_histogram(binwidth = 10, position = "dodge") +
  labs(title = "Monthly Charges Distribution by Customer Status")

```
## Observations :  
- Tenure: Customers who left tend to have lower tenure.

- MonthlyCharges: Higher charges are linked to higher churn.

```{r}
# Contract type vs churn
ggplot(Model_Building_Data, aes(x = Contract, fill = Status)) +
  geom_bar(position = "fill") +
  labs(title = "Contract Type vs Customer Status", y = "Proportion")

# Payment method vs churn
ggplot(Model_Building_Data, aes(x = PaymentMethod, fill = Status)) +
  geom_bar(position = "fill") +
  labs(title = "Payment Method vs Customer Status", y = "Proportion") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
## Observations :  
- Customers with month-to-month contracts have higher churn.

- Electronic check is associated with higher churn.

- Longer contracts (1 or 2 years) see more retention.

```{r}
set.seed(123)
train_index <- createDataPartition(Model_Building_Data$Status, p = 0.8, list = FALSE)

train_df <- Model_Building_Data[train_index, ]
test_df <- Model_Building_Data[-train_index, ]


```


## Question - 2
```{r}
library(tidyverse)
library(caret)
library(pROC)

# Load and clean data
model_df <- read.csv("Model_Building_Data.csv")
model_df <- model_df[!is.na(model_df$TotalCharges), ]

# Convert Status to factor with correct reference level
model_df$Status <- factor(model_df$Status, levels = c("Current", "Left"))

# Train-test split (80/20)
set.seed(123)
train_index <- createDataPartition(model_df$Status, p = 0.8, list = FALSE)
train_df <- model_df[train_index, ]
test_df <- model_df[-train_index, ]

# Logistic regression model (using all variables)
log_model <- glm(Status ~ Gender + SeniorCitizen + Partner + Dependents +
                   Tenure + PhoneService + MultipleLines + InternetService +
                   OnlineSecurity + DeviceProtection + StreamingMovies +
                   Contract + PaperlessBilling + PaymentMethod +
                   MonthlyCharges + TotalCharges,
                 data = train_df, family = "binomial")

# Model summary
summary(log_model)

# Predict probabilities on test data
logistic_prob <- predict(log_model, newdata = test_df, type = "response")

# ROC and AUC
roc_obj <- roc(test_df$Status, logistic_prob)
auc(roc_obj)
plot(roc_obj, main = "ROC Curve for Logistic Regression", col = "blue")
```
# Logistic Regression Coefficient Interpretations :  
- Tenure (Estimate = -0.06398)
- A one-unit increase in tenure (i.e., staying one additional month) reduces the odds of churn by a factor of exp(-0.06398), holding other variables constant.  

- TotalCharges (Estimate = 0.0003647)
- A one-unit increase in total charges is associated with a very small increase in the odds of churn by a factor of exp(0.0003647).    

- SeniorCitizen (Estimate = 0.3056)
- Being a senior citizen increases the odds of churn by a factor of exp(0.3056) compared to a non-senior, holding all else equal.  

- PhoneServiceYes (Estimate = -0.5431)
- Customers with phone service have lower odds of churn by a factor of exp(-0.5431) compared to those without.  

- MultipleLinesYes (Estimate = 0.2334)
- Having multiple lines increases the odds of churn by a factor of exp(0.2334)  compared to having a single line.  

- InternetServiceFiber optic (Estimate = 1.107)
- Customers with fiber optic internet have much higher odds of churn — by a factor of exp(1.107)  compared to those with DSL.  

- InternetServiceNo (Estimate = -0.8469)
- Customers without internet service have much lower odds of churn — by a factor of exp(-0.8469).  

- OnlineSecurityYes (Estimate = -0.3775)
- Customers who have online security are less likely to churn, with odds reduced by a factor of exp(-0.3775.  

- StreamingMoviesYes (Estimate = 0.2853)
- Customers who stream movies are more likely to churn, with odds increased by a factor of exp(0.2853). 

- ContractOne year (Estimate = -0.7025)
- Customers with a one-year contract have lower odds of churn — by a factor of exp(-0.7025) compared to those with month-to-month plans.  
- ContractTwo year (Estimate = -1.633)
- Customers on two-year contracts have much lower odds of churn — by a factor of exp(-1.633).  

- PaymentMethodElectronic check (Estimate = 0.3498)
- Customers who pay via electronic check are more likely to churn, with odds increased by a factor of exp(0.3498).  

# Classification using fixed threshold of 0.10 
```{r}
logistic_class_0.10 <- ifelse(logistic_prob > 0.10, "Left", "Current")
table(logistic_class_0.10, test_df$Status)

```
# Calculating the Prediction Accuracy  
```{r}
(471+324)/(471+22+480+324)
```

# Classification using fixed threshold of 0.15  
```{r}
logistic_class_0.15 <- ifelse(logistic_prob > 0.15, "Left", "Current")
table(logistic_class_0.15, test_df$Status)
```
# Calculating the Prediction Accuracy  
```{r}
(546+316)/(546+30+405+316)
```
# Classification using fixed threshold of 0.05  

```{r}
logistic_class_0.05 <- ifelse(logistic_prob > 0.05, "Left", "Current")
table(logistic_class_0.05, test_df$Status)

```

# Calculating the Prediction Accuracy  
```{r}
(352+336)/(352+10+599+336)
```

## Observations :   
# Model Performance :  
- The AUC value of the logistic regression model is approximately r log_auc, indicating good model performance in distinguishing between churned and retained customers.

- Using a threshold of 0.10, we predicted churn effectively, capturing more churners (sensitivity) while accepting a few false positives.  

# Coefficient Interpretations:  
- Contract type has a strong influence on churn. Customers with two-year contracts are significantly less likely to leave, while those on month-to-month contracts have a high likelihood of churn.

- Customers paying via Electronic Check are more likely to churn compared to those using credit cards or bank transfers.

- Monthly Charges is positively associated with churn – higher bills correlate with higher likelihood of leaving.  
## third question  

```{r}
library(e1071)
library(pROC)
```

```{r}
# Model 1: All variables
nb_model1 <- naiveBayes(Status ~ ., data = train_df)

# Model 2: Selected predictors based on earlier insights
nb_model2 <- naiveBayes(Status ~ Contract + MonthlyCharges + Tenure + OnlineSecurity + PaymentMethod + TotalCharges, data = train_df)

# Model 3: Minimal predictors
nb_model3 <- naiveBayes(Status ~ Contract + MonthlyCharges + Tenure, data = train_df)

```

```{r}
# Model 1
nb_prob1 <- predict(nb_model1, newdata = test_df, type = "raw")[, "Left"]

# Model 2
nb_prob2 <- predict(nb_model2, newdata = test_df, type = "raw")[, "Left"]

# Model 3
nb_prob3 <- predict(nb_model3, newdata = test_df, type = "raw")[, "Left"]


```

```{r}
# AUC for every model
nb_roc1 <- roc(test_df$Status, nb_prob1)
nb_roc2 <- roc(test_df$Status, nb_prob2)
nb_roc3 <- roc(test_df$Status, nb_prob3)

auc(nb_roc1)
auc(nb_roc2)
auc(nb_roc3)


```

```{r}
# Plotting ROC for best model 
plot(nb_roc2, main = "ROC Curve - Naive Bayes (Best Model)", col = "darkgreen")


```

```{r}

# Threshold 0.10
nb_class_0.10 <- ifelse(nb_prob2 > 0.10, "Left", "Current")
table(nb_class_0.10, test_df$Status)

# Threshold 0.15
nb_class_0.15 <- ifelse(nb_prob2 > 0.15, "Left", "Current")
table(nb_class_0.15, test_df$Status)

# Threshold 0.50
nb_class_0.50 <- ifelse(nb_prob2 > 0.50, "Left", "Current")
table(nb_class_0.50, test_df$Status)



```
## observations :  
# Model Comparison:  
- Among the three Naive Bayes models tested, the one using predictors like Contract, Tenure, MonthlyCharges,  OnlineSecurity, PaymentMethod,  and TotalCharges  had the highest AUC of approximately r nb_auc.  

- Naive Bayes assumes feature independence, which is a limitation, but it still performs well with carefully chosen variables.   

# Threshold Evaluation:  
- At a 0.10 threshold, the model successfully identifies customers with a higher risk of leaving, providing a strong base for proactive business intervention.

- Although false positives exist, the cost of losing a customer is usually higher than the cost of a false alarm, so this model is practical for early warnings.


## question - 4

```{r}
# Model 1: All predictors
library(MASS)

lda_model1 <- lda(Status ~ Gender + SeniorCitizen + Partner + Dependents +
                    Tenure + PhoneService + MultipleLines + InternetService +
                    OnlineSecurity + DeviceProtection + StreamingMovies +
                    Contract + PaperlessBilling + PaymentMethod +
                    MonthlyCharges + TotalCharges,
                  data = train_df)

# Model 2: Based on strongest predictors from previous models
lda_model2 <- lda(Status ~ Contract + MonthlyCharges + Tenure + PaymentMethod + OnlineSecurity + TotalCharges,
                  data = train_df)

# Model 3: Minimal model
lda_model3 <- lda(Status ~ Contract + MonthlyCharges + Tenure,
                  data = train_df)


```

```{r}
# Model 1
lda_pred1 <- predict(lda_model1, newdata = test_df)
lda_prob1 <- lda_pred1$posterior[, "Left"]

# Model 2
lda_pred2 <- predict(lda_model2, newdata = test_df)
lda_prob2 <- lda_pred2$posterior[, "Left"]

# Model 3
lda_pred3 <- predict(lda_model3, newdata = test_df)
lda_prob3 <- lda_pred3$posterior[, "Left"]

```


```{r}
# AUC
lda_roc1 <- roc(test_df$Status, lda_prob1)
lda_roc2 <- roc(test_df$Status, lda_prob2)
lda_roc3 <- roc(test_df$Status, lda_prob3)

auc(lda_roc1)
auc(lda_roc2)
auc(lda_roc3)

```


```{r}
# Plot ROC for best LDA model (Model 3)
plot(lda_roc3, main = "ROC Curve - LDA (Best Model)", col = "purple")

```


```{r}

# Threshold 0.10
lda_probability3 <- lda_prob3  # just to use your naming
LDA_test_data2 <- test_df      # your variable naming style
lda_class_0.10 <- ifelse(lda_probability3 > 0.10, "Left", "Current")
table(lda_class_0.10, LDA_test_data2$Status)

# Threshold 0.15
lda_class_0.15 <- ifelse(lda_probability3 > 0.15, "Left", "Current")
table(lda_class_0.15, LDA_test_data2$Status)

# Threshold 0.50
lda_class_0.50 <- ifelse(lda_probability3 > 0.50, "Left", "Current")
table(lda_class_0.50, LDA_test_data2$Status)


```
## Observations :  
# Model Performance:  
- The best-performing LDA model used Contract, MonthlyCharges, and Tenure as predictors, achieving an AUC of approximately r lda_auc.

- LDA performed well with these selected features, indicating that customer contract type and billing level provide clear separation between churned and retained customers.  

## Threshold Evaluation:  
- Using a 0.10 threshold helped identify more potential churners, prioritizing sensitivity.  

- The confusion matrix revealed a reasonable trade-off between capturing churners and avoiding too many false positives.  

- LDA also provides posterior probabilities, making it useful for ranking customers by churn risk.  
 
## question - 5

```{r}
library(MASS)
qda_model_simple <- qda(Status ~ Contract + Tenure + MonthlyCharges + TotalCharges,
                        data = train_df)

# Predict probabilities
qda_pred_simple <- predict(qda_model_simple, newdata = test_df)
qda_prob_simple <- qda_pred_simple$posterior[, "Left"]


```


```{r}
qda_probability <- qda_prob_simple
qda_test_data <- test_df

# Threshold 0.10
qda_class_0.10 <- ifelse(qda_probability > 0.10, "Left", "Current")
table(qda_class_0.10, qda_test_data$Status)

# Threshold 0.15
qda_class_0.15 <- ifelse(qda_probability > 0.15, "Left", "Current")
table(qda_class_0.15, qda_test_data$Status)

# Threshold 0.50
qda_class_0.50 <- ifelse(qda_probability > 0.50, "Left", "Current")
table(qda_class_0.50, qda_test_data$Status)

```


```{r}
library(pROC)

qda_roc_simple <- roc(qda_test_data$Status, qda_probability)
auc(qda_roc_simple)
plot(qda_roc_simple, main = "ROC Curve - QDA (Simplified Model)", col = "brown")

```
## Observations :  
# Model Adjustment:  
- A full QDA model failed due to rank deficiency, which is common when predictors are highly correlated or when there are more predictors than needed.  

- A simplified model using only numeric variables like Contract, Tenure, MonthlyCharges, and TotalCharges was successful and had a competitive AUC of approximately r qda_auc.  

# Threshold Evaluation:  
- Like LDA, QDA also worked best with a 0.10 threshold, detecting many at-risk customers.

- It offers flexibility in capturing nonlinear boundaries, which may help in capturing subtle churn patterns.

## question - 6

```{r}
library(tree)
library(caret)

# Fitting tree models with different predictor combinations

# Model 1: All predictors
tree_model1 <- tree(Status ~ ., data = train_df)

# Model 2: Selected predictors
tree_model2 <- tree(Status ~ Contract + Tenure + MonthlyCharges + OnlineSecurity + PaymentMethod,
                    data = train_df)

# Model 3: Minimal predictors
tree_model3 <- tree(Status ~ Contract + MonthlyCharges + Tenure,
                    data = train_df)

```


```{r}
# Model 1
tree_pred1 <- predict(tree_model1, newdata = test_df, type = "class")
tree_pred1 <- factor(tree_pred1, levels = c("Current", "Left"))
confusionMatrix(tree_pred1, factor(test_df$Status, levels = c("Current", "Left")))

# Model 2
tree_pred2 <- predict(tree_model2, newdata = test_df, type = "class")
tree_pred2 <- factor(tree_pred2, levels = c("Current", "Left"))
confusionMatrix(tree_pred2, factor(test_df$Status, levels = c("Current", "Left")))

# Model 3
tree_pred3 <- predict(tree_model3, newdata = test_df, type = "class")
tree_pred3 <- factor(tree_pred3, levels = c("Current", "Left"))
confusionMatrix(tree_pred3, factor(test_df$Status, levels = c("Current", "Left")))


```

```{r}
# Plot the trees
plot(tree_model1); text(tree_model1, pretty = 0, cex = 0.7, main = "Decision Tree - Model 1")
plot(tree_model2); text(tree_model2, pretty = 0, cex = 0.7, main = "Decision Tree - Model 2")
plot(tree_model3); text(tree_model3, pretty = 0, cex = 0.7, main = "Decision Tree - Model 3")


```
## 0bservations:  
# Model Comparison:  
- Among the three tree models, the one using Contract, Tenure, MonthlyCharges, OnlineSecurity, and PaymentMethod gave the highest prediction accuracy on the test set.  

- While decision trees don’t provide probabilities or AUC like other models, they are highly interpretable.  

# Business Utility:  
- The tree structure clearly highlighted that customers on month-to-month contracts, with no online security, and higher monthly charges are at the greatest risk.

- These insights are easy to communicate to business teams and ideal for rule-based interventions.  

## question - 7

```{r}
library(pROC)
library(caret)


```


```{r}
# AUC for Logistic Regression
log_auc <- auc(test_df$Status, logistic_prob)

# AUC for Naive Bayes (Model 2)
nb_auc <- auc(test_df$Status, nb_prob2)

# AUC for LDA (Model 3)
lda_auc <- auc(test_df$Status, lda_prob3)

# AUC for QDA (simplified model)
qda_auc <- auc(qda_test_data$Status, qda_probability)

# Print all AUCs
log_auc
nb_auc
lda_auc
qda_auc

```


```{r}
# Logistic Regression (threshold 0.10)
logistic_class_0.10 <- ifelse(logistic_prob > 0.10, "Left", "Current")
log_cm <- table(logistic_class_0.10, test_df$Status)
log_accuracy <- sum(diag(log_cm)) / sum(log_cm)

# Naive Bayes (threshold 0.10)
nb_class_0.10 <- ifelse(nb_prob2 > 0.10, "Left", "Current")
nb_cm <- table(nb_class_0.10, test_df$Status)
nb_accuracy <- sum(diag(nb_cm)) / sum(nb_cm)

# LDA (threshold 0.10)
lda_class_0.10 <- ifelse(lda_prob3 > 0.10, "Left", "Current")
lda_cm <- table(lda_class_0.10, test_df$Status)
lda_accuracy <- sum(diag(lda_cm)) / sum(lda_cm)

# QDA (threshold 0.10)
qda_class_0.10 <- ifelse(qda_probability > 0.10, "Left", "Current")
qda_cm <- table(qda_class_0.10, qda_test_data$Status)
qda_accuracy <- sum(diag(qda_cm)) / sum(qda_cm)

# Decision Tree (assume tree_model2 was best)
tree_pred <- predict(tree_model2, newdata = test_df, type = "class")
tree_pred <- factor(tree_pred, levels = c("Current", "Left"))
tree_cm <- table(tree_pred, test_df$Status)
tree_accuracy <- sum(diag(tree_cm)) / sum(tree_cm)

# Print all accuracies
log_accuracy
nb_accuracy
lda_accuracy
qda_accuracy
tree_accuracy

```
## Observations :  

# AUC Comparison:  
- Logistic Regression had the highest AUC of approximately r log_auc, indicating it is the best at separating churned from current customers.  

- Naive Bayes and LDA followed closely with slightly lower AUCs.  

- QDA, even with a simplified model, performed competitively but did not outperform logistic regression.  

# Accuracy Comparison (Threshold = 0.10):  
- Decision Tree model had the highest prediction accuracy on the test set, based on the confusion matrix.  

- While it performed best in raw accuracy, it lacks the probabilistic interpretation provided by the other models.  

# Conclusion :  
- Logistic regression is the best overall model when combining AUC and interpretability.  

- Decision trees are useful for business rule generation but may not be as flexible or robust for probability-based risk scoring.  

## question - 8 

```{r}
# Load the implementation data
implementation_df <- read.csv("Implementation_Data.csv")

# Predict churn probabilities using logistic regression model
logistic_pred_impl <- predict(log_model, newdata = implementation_df, type = "response")

```


```{r}
# Classify customers using threshold 0.10
implementation_df$Churn_Prediction <- ifelse(logistic_pred_impl > 0.10, "Left", "Current")

# Count predicted churners
table(implementation_df$Churn_Prediction)

```


```{r}
# Total revenue from predicted churners
churned_customers <- implementation_df[implementation_df$Churn_Prediction == "Left", ]
revenue_loss <- sum(churned_customers$MonthlyCharges)
revenue_loss

```


```{r}
# Sort coefficients by absolute value 
log_coeff <- summary(log_model)$coefficients
sorted_importance <- sort(abs(log_coeff[-1, 1]), decreasing = TRUE)
sorted_importance

```


```{r}
# Set parameters
credit_offer <- 20
retention_rate <- 0.5

# Estimating how many customers will stay
retained_customers <- nrow(churned_customers) * retention_rate

# Total revenue retained
retained_revenue <- sum(churned_customers$MonthlyCharges) * retention_rate

# Total incentive cost
incentive_cost <- credit_offer * retained_customers

# Net benefit
net_benefit <- retained_revenue - incentive_cost
net_benefit

```
## Observations :  
# Churn Prediction:  
- The logistic regression model flagged approximately r nrow(churned_customers) customers as likely to churn using a 0.10 threshold.  

- These customers account for a potential monthly revenue loss of around r revenue_loss if they leave.  
# Proposed Incentive Plan:  
- Offering a $20/month credit could retain 50% of the at-risk customers.

- Based on the simulation:  

- Estimated revenue retained: r retained_revenue

- Estimated incentive cost: r incentive_cost

- Net monthly benefit: r net_benefit

# Final Recommendation:  
- Launch a targeted retention campaign offering credits or perks to high-risk customers.  

- Focus on customers with month-to-month contracts, no online security, and high monthly charges.  

- This strategy can yield a positive return while reducing churn and increasing customer satisfaction.  



